---
title: "ETV6 Fusions in Pediatric AML"
author: "Jenny Smith"
date: "8/12/20"
output: html_document
---


#Set-up

```{r setup, cache = FALSE, include = FALSE}
require(knitr)
knitr::opts_knit$set(root.dir = file.path(PROJHOME,"2020.08.12_ETV6_Fusions"))
```

```{r}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),
                      tidy=TRUE,
                      fig.align='center',
                      fig.width = 10, fig.height = 10)
# options(stringsAsFactors = FALSE,bitmapType = 'cairo')
# grDevices::X11.options(type='cairo')
```

```{r message=FALSE}
library(ggplot2)
library(ggrepel)
library(gridExtra)
library(RColorBrewer)

library(readr)
library(dplyr)
library(magrittr)
library(tibble)
library(tidyr)
library(data.table)
library(stringr)
library(readr)


library(tools)
library(gtools)

# library(DeGSEA)
getwd()
```


#Read in the counts data

```{r}
cts <- readRDS(file.path(PROJHOME, "0000.00.03_ExpressionMatrices/TARGET_AML_DSAML_MPN_NBM_Ribodepleted_dupGenesRemoved_Fractionalcounts.RDS"))

dim(cts) #48230  2345 (filtered to remove all zero count rows)
```

```{r}
cts <- as.matrix(cts[rowSums(cts)>= 10, ]) 

dim(cts) #48230  2345
```

```{r}
TPM <- readRDS(file.path(PROJHOME, "0000.00.03_ExpressionMatrices/TARGET_AML_DSAML_MPN_NBM_Ribodepleted_dupGenesRemoved_TPM.RDS"))

dim(TPM) #51573  2345 (filtered to remove all zero count rows)
```



#Read in the Clinical Data

```{r}
CDEs <- read.csv(file.path(CDE,"Merged/TARGET_AML_0531_1031_merged_CDEs_7.08.20.csv"),
                   na.strings = c("N/A","#N/A","NA","^$", "^\\.$"))

CDEs <- CDEs %>% 
  filter(!is.na(USI) | USI != "Unknown")

dim(CDEs) #2396  143
```

```{r}
manifest <- read.csv(file.path(TARGET,"SequencingDataMatrix/TARGET_AML_Ribodepleted_Master_Manifest_8.5.20.csv"), 
                     row.names = 1)

head(manifest[,1:5])
dim(manifest) #2346   14
```


#Read in the fusion data 

```{r eval=FALSE}
fusions <- read.csv(file.path(PROJHOME,"2018.09.11_Combine_Fusion_Calls/Combined_withConfidence_Intervals/TARGET_AML_0531_1031_Relapse_Combined_STAR_TransAbyss_CICERO_FusionCalls_withConfidenceLevels_Annotated_6.17.2020.csv"))

dim(fusions)
```

```{r}
curated_fusions <- read.csv(file.path(PROJHOME,"2018.09.11_Combine_Fusion_Calls/Combined_withConfidence_Intervals/TARGET_AML_0531_1031_Relapse_Combined_STAR_TransAbyss_CICERO_Cleaned_Annotated_FOI_FusionCalls_6.17.20.csv"))


head(curated_fusions[,1:5])
dim(curated_fusions)
```

#Check for discrepancies 
Ideally, this will not be standard practice. Need to include the ~40 patients who were re-categorized to a different fusion in the final working CDEs.

```{r}
check <- filter(curated_fusions, !is.na(Updated_Primary.Fusion)) %>% 
  filter(grepl("ETV6",Updated_Primary.Fusion))


# View(check)
dim(check) #OK none were changed to a different fusion or found to be false positive s
```



#Define Samples

Jenny, can you provide me this information?
1.  Incidence of ETV6 fusions in all patients
2.  Incidence of ETV6 in those < 1 year
3.  Incidence of ETV6 fusion in those <2 years
4.  number of different ETV6 fusions (provide number for each fusion partner)

```{r}
ETV6.all <- CDEs %>% 
  mutate(ETV6_Fusions=case_when(
    grepl("ETV6", Primary.Fusion.CNV) | grepl("ETV6", Additional.Fusions.CNV) ~ "ETV6", 
    grepl("No", ScreenedForFusion) ~ "Unknown",
    TRUE ~ "OtherAML")) %>%
  mutate_at(vars(ETV6_Fusions),
            ~ifelse(grepl("ETV6 (deletion)",Primary.Fusion.CNV, fixed = T) | 
                      grepl("ETV6 (deletion)",Additional.Fusions.CNV, fixed = T), "OtherAML", .)) %>%
  mutate(Age_Under1yr=case_when(
    Age.in.years < 1.0 ~ "Less than  1 yr",
    grepl("Unknown",Age.Category) ~ "Unknown",
    TRUE ~ "Greater than 1 yr")) %>%
  mutate(Age_Under2yr=case_when(
    Age.in.years < 2.0 ~ "Less than 2 yrs",
    grepl("Unknown",Age.Category) ~ "Unknown",
    TRUE ~ "Greater than 2 yrs")) %>%
  filter(ETV6_Fusions != "Unknown")


table(ETV6.all$ETV6_Fusions, useNA='ifany')
table(ETV6.all$Age_Under1yr, useNA='ifany')
table(ETV6.all$Age_Under2yr, useNA='ifany')
```


Three patients with ETV6 deletions: 
	
PAVKVT	ETV6 (deletion)
PAWHIZ	PICALM-MLLT10	ETV6 (deletion)	
PAXARD	KMT2A-MLLT3	ETV6 (deletion)	

```{r}
samples_Check <- ETV6.all %>%
  select(USI,Protocol, ISCN,Primary.Fusion.CNV,Additional.Fusions.CNV,
       ScreenedForFusion_byNGS=ScreenedForFusion,Cyto.vs..Seq,
       ETV6_Fusions,) %>% 
       filter(ETV6_Fusions == "ETV6") 

# write.csv(samples_Check, "ETV6_Fusion_Positive_Patients.csv",row.names = FALSE)
```

```{r}
filter(curated_fusions, grepl("PARJGD|PAUPDK", USI))
```



#Frequencies of ETV6 Fusions

```{r}
ETV6_tab1 <- ETV6.all %>% 
  mutate(Total=n()) %>%
  group_by(ETV6_Fusions, Total) %>% 
  summarise(Number_of_Samples=n()) %>% 
  ungroup() %>%
  mutate(Percent_of_Samples=round(Number_of_Samples/Total *100, digits = 2))

ETV6_tab1
# write.csv(ETV6_tab1, "TARGET_AML_ETV6_Fusion_Overall_Frequency.csv", row.names = FALSE)
```

```{r}
table(ETV6.all$ETV6_Fusions, ETV6.all$Age_Under1yr)
```


#Incidence by Age

for the incidence, can you provide the flip side of the data you provided.  as in
1.  in all patients that are <1 years of age (N=??), how many (N=??; ??%) are ETV6 fusion positive.
2.   in all patients that are <2 years of age (N=??), how many (N=??; ??%) are ETV6 fusion positive.


```{r}
ETV6_1yr <- ETV6.all %>% 
  mutate(Total=n()) %>%
  group_by(Age_Under1yr, Total) %>%
  mutate(Number_Per_Age_Group=n())  %>%
  
  
  group_by(ETV6_Fusions,Number_Per_Age_Group, add=TRUE) %>%
  summarise(Number_of_Samples=n()) %>%
  ungroup() %>%
  
  group_by(Age_Under1yr) %>%
  mutate(Percent_of_Samples_byAge_Group=round(Number_of_Samples/sum(Number_of_Samples) *100, digits = 2)) %>%
  ungroup()

  # group_by(ETV6_Fusions) %>%
  # mutate(Percent_of_Samples_byETV6_Group=round(Number_of_Samples/sum(Number_of_Samples) *100, digits = 2)) %>%
  # ungroup()

ETV6_1yr
# write.csv(ETV6_1yr, "TARGET_AML_ETV6_Fusion_Frequency_1yrOldComparison.csv")
```

```{r}
ETV6_2yr  <- ETV6.all %>% 
  mutate(Total=n()) %>%
  group_by(Age_Under2yr, Total) %>%
  mutate(Number_Per_Age_Group=n())  %>%
  
  
  group_by(ETV6_Fusions,Number_Per_Age_Group, add=TRUE) %>%
  summarise(Number_of_Samples=n()) %>%
  ungroup() %>%
  
  group_by(Age_Under2yr) %>%
  mutate(Percent_of_Samples_byAge_Group=round(Number_of_Samples/sum(Number_of_Samples) *100, digits = 2)) %>%
  ungroup()

  
  # group_by(ETV6_Fusions) %>%
  # mutate(Percent_of_Samples_byETV6_Group=round(Number_of_Samples/sum(Number_of_Samples) *100, digits = 2)) %>%
  # ungroup()

ETV6_2yr
# write.csv(ETV6_2yr, "TARGET_AML_ETV6_Fusion_Frequency_2yrOld_Comparison.csv")
```

#Summarize Fusions

```{r}
ETV6_fusions_freq <- ETV6.all %>% 
  filter(ETV6_Fusions == "ETV6") %>%
  mutate(Fusion=case_when(
    ! grepl("ETV6", Primary.Fusion.CNV) ~ Additional.Fusions.CNV,
    TRUE ~ Primary.Fusion.CNV)) %>% 
  mutate(Fusion=ifelse(grepl("NIPBL-ETV6", Fusion), "ETV6-NIPBL", Fusion)) %>%
  select(USI,ETV6_Fusions, Fusion) %>%
  
  mutate(Total_ETV6_Fusions=n()) %>%
  group_by(Fusion,Total_ETV6_Fusions) %>% 
  summarise(Frequency=n()) %>% 
  ungroup() %>% 
  mutate(Percent=round(Frequency/Total_ETV6_Fusions*100, digits = 2)) %>%
  arrange(desc(Frequency))

ETV6_fusions_freq
# View(ETV6_fusions_freq)

# write.csv(ETV6_fusions_freq, "TARGET_AML_ETV6_Fusions_Frequency.csv",row.names = FALSE)
```




#Session Information

```{r}
sessionInfo()
```

